---
title: "Scientific Worflows and Good Pratices"
author: "Tawane Nunes"
date: '2022-07-07'
output: html_document
---

# Scientific workflow

>Reproducibility - for you, collegues, and community
>Script based tools
>Version control
>Share methods
>Peer review

__Never create one single script with all the analysis__ 

Ideally each scripts stars reading a particular input/data and ends writing results
The first script will read and format the data, and the others have specific analisys. Needs to run the first one.

It's advisable to never sabe the workspace, but if you have an object that thakes a lot to run you can save as _.rda_ using _save()_ and to load the object after you can use _load()_.

#Each script

1. A header containing whi, how, when, where, and why METADA

2. Loading all package

3. Reads needed data

4. Coding a variable that will not change

5. Is good to comment all steps

6. Writing in the HD the result from each step

7. The sxript must be able to be run in sequence from start to finish. (no repetitions, no lines out of order, no parenteheses or non-closing calls)

8. You should be able to erase the workspace mid-session and rebuild

9. Do not define functions inside script

10. use concise and informative names and don't use names already taken

11. If you copy and paste more than three times it's time to write a loop or a function

# Hands on

Data of plant community

Listing all files in the path and saving in an object

```{r}
cestes_files <- list.files(path = "data/raw/cestes",
                           pattern = "csv$",
                           full.names = TRUE)
```

Getting the names of the arquives automatically without .csv using __gsub()__

```{r}
cestes_names <- gsub(".csv", "", basename(cestes_files), fixed = TRUE)

envir <- read.csv(cestes_files[3])
```

Creating a loop to read all the csv files

```{r}

data_list <- lapply(cestes_files, read.csv)

#adding names with the vector created before with sfub
names(data_list) <- cestes_names
```

Inspecting the data 

```{r}
length(data_list)

# Inspecting data
head(data_list$envir)
dim(data_list$envir)
summary(data_list$envir)
```

Creating summary table for all environmetal variables

```{r}
sd(envir$Clay)
envir_mean <- apply(envir[, -1], 2, mean) # mean of each variable
envir_sd <- apply(envir[, -1], 2, sd) # sd of each variable

```

Creating a function to standard deviation and if round is true it will round

```{r}
std <- function(x, round = FALSE, ...) { # the tree dots you can choose arguments for the function that is inside the function, in this case is round
  std <- sd(x) / sqrt(length(x))
  if (round) std <- round(std, ...) # here the three dots indicate that the argument will be defined in the major function
  return(std)
}
```


