---
title: "Time-series"
description: |
author:
  - name: Sara Mortara
    affiliation: re.green | ¡liibre!
    url: {}
    orcid_id: 0000-0001-6221-7537
  - name: Andrea Sánchez-Tapia
    orcid_id: 0000-0002-3521-4338
    affiliation: ¡liibre!
    affiliation_url: {}
    url: https://andreasancheztapia.netlify.app
date: "2022-07-21"
output:
  distill::distill_article:
    toc: true
    theme: theme.css
---

In this exercise we will learn how to manipulate dates in R using the package __lubridate__. We will also use the package __zoo__ to calculate the rolling mean of a variable.

```{r}
library(dplyr)
library(ggplot2)
library(lubridate)
library(zoo)
```

In our example, we will use Covid-19 data from Recife in Pernambuco, Brazil. 

```{r, echo=FALSE, fig.align='center'}
library(sf)
library(tmap)
br <- brazilmaps::get_brmap(geo = "State")
pe <- br %>% filter(nome == "PERNAMBUCO")
tm_shape(br) + tm_polygons() +
  tm_shape(pe) + tm_polygons(col = "darkgrey")

```

This data was downloaded from the portal [Brasil.IO](https://brasil.io/). 

```{r}
covid <- read.csv("data/raw/covid19-dd7bc8e57412439098d9b25129ae6f35.csv")

```

## Converting into date format

```{r}
# First checking the class
class(covid$date)
# Changing to date format
covid$date <- as_date(covid$date)
# Checking the class
class(covid$date)
# Now we can make numeric operations
range(covid$date)
```

## Plotting a time-series with __ggplot2__

First, we will create a column containing the number of new cases. 


```{r}
ggplot(covid) +
  geom_line(aes(x = date, y = new_confirmed)) +
  theme_minimal()
```

Oops. We have negative cases and will substitute the negative values per zero. 

```{r}
covid$new_confirmed[covid$new_confirmed < 0] <- 0 
```

Let's try again.

```{r}
p <- ggplot(covid) +
  geom_line(aes(x = date, y = new_confirmed), color = "grey30") +
  labs(x = "Date", y = "New cases") +
  theme_minimal()
p
```

We can change the date axis in ggplot using the function `scale_x_date()`.

```{r}
p + scale_x_date(date_breaks = "4 months", date_labels = "%b-%y")

```

## Rolling mean with __zoo__ package

We will use the function `rollmean()` from the __zoo__ package to calculate the rolling mean of the new cases for a 14 day interval.

```{r}
# Rolling mean
covid$new_rollmean <- rollmean(x = covid$new_confirmed, k = 14, align = "right", 
                               fill = NA)
```

Now, we will add a red line with the rolling mean. 

```{r}
p + scale_x_date(date_breaks = "4 months", date_labels = "%b-%y") +
  geom_line(data = covid, aes(x = date, y = new_rollmean), color = "red")
```

